---
title: "Part 4"
author: "Claire Madden"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
#Load packages

library(tidyverse)
library(stargazer)
library(car)

#Load edited dataset

faculty_salaries <- read_csv("faculty data edited.csv")
```

```{r, include = FALSE}
#Some initial data exploration to get an early sense of any trends

#Salary by gender
sal_by_gender <- faculty_salaries %>% 
  group_by(Sex) %>% 
  summarize(
    mean = mean(Salary)
  )

#Salary by rank (professor, associate professor, assistant professor)
sal_by_rank <- faculty_salaries %>% 
  group_by(Rank) %>% 
  summarize(
    mean = mean(Salary)
  )

#Average years (since PhD, and of service) for each rank
sal_by_bothyears <- faculty_salaries %>% 
  group_by(Rank) %>% 
  summarize(
    mean_PhD = mean(Years_PhD),
    mean_service = mean(Years_Service)
  )

#Salary by years of service (number of years as faculty), color coded for gender, separated by discipline
sal_by_service <- ggplot(faculty_salaries, aes(x = Years_Service, y = Salary)) +
  geom_point(aes(color = Sex), alpha = 0.5) +
  facet_wrap(~ Discipline)

sal_by_service

#Salary by years since PhD, color coded for gender, separated by discipline
sal_by_years <- ggplot(faculty_salaries, aes(x = Years_PhD, y = Salary)) +
  geom_point(aes(color = Sex), alpha = 0.5) +
  facet_wrap(~ Discipline)

sal_by_years

#Salary by years of service (number of years as faculty), color coded for gender, separated by rank
sal_by_service_rank <- ggplot(faculty_salaries, aes(x = Years_Service, y = Salary)) +
  geom_point(aes(color = Sex), alpha = 0.5) +
  facet_wrap(~ Rank)

sal_by_service_rank

#Salary by years since PhD, color coded for gender, separated by rank
sal_by_years_rank <- ggplot(faculty_salaries, aes(x = Years_PhD, y = Salary)) +
  geom_point(aes(color = Sex), alpha = 0.5) +
  facet_wrap(~ Rank)

sal_by_years_rank

#Mean salary by discipline (A = theoretical, B = applied)
sal_discipline_mean <- faculty_salaries %>% 
  group_by(Discipline) %>% 
  summarize(
    mean = mean(Salary)
  )

```

```{r, include = FALSE}
#Exploring possibilities of collinearity (focusing on years PhD, years of service)

col_df <- faculty_salaries %>% 
  select(Years_PhD, Years_Service, Salary)

pairs(col_df)

#Years_PhD and Years_Service appear to be highly correlated

```


```{r, include = FALSE}
#Creating a saturated model for salary, including all variables:

#Rank - professor, associate professor, assistant professor
#Discipline - A = theoretical or B = applied
#Sex - male or female
#Years_PhD - years since getting PhD
#Years_Service - years as a faculty member

sal_lm1 <- lm(Salary ~ Rank + Discipline + Sex + Years_PhD + Years_Service, data = faculty_salaries)

summary(sal_lm1)

#Info from the model summary:
#For rank, the referene level is associate professor
#For sex, the reference level is female
#For discipline, the reference level is A (theoretical)
#Years_PhD and Years_Service are continuous 

#Coefficients for rank seem to make sense - compared to associate professor, a full professor makes a good deal more, and an assistant professor makes less (matches the means that were calculated above)

#Coefficient for sex seems to make sense - male faculty made more

#Coefficient for discipline seems to make sense - applied made more than theoretical

#Are years_PhD and years_service redundant? Do they sort of get at a similar thing? Which one is better? Are they both incorporated within rank?

#Based on overall p-value, the model does significantly predict salary (p < 0.001)

```

```{r, include = FALSE}
#Another model, this time using years_service and not years_phd

sal_lm2 <- lm(Salary ~ Rank + Discipline + Sex + Years_Service, data = faculty_salaries)

summary(sal_lm2)

#Still weird that there is a negative coefficient for years of service - doesn't make sense!
```

```{r, include = FALSE}
#What about a model with years since phd, instead of years of service?

sal_lm3 <- lm(Salary ~ Rank + Discipline + Sex + Years_PhD, data = faculty_salaries)

summary(sal_lm3)

#Positive coefficient for years since PhD, so better!
```

```{r, include = FALSE}
#Now a model with only rank, discipline, and sex (not using years since PhD or years of service)

sal_lm4 <- lm(Salary ~ Rank + Discipline + Sex, data = faculty_salaries)

summary(sal_lm4)
```


```{r, include = FALSE}
#Model diagnostics

#For lm1 (saturated model)

plot(sal_lm1)

#Homoscedasticity seems fine, residuals normality looks pretty good for majority of data (upper tail is a little strange)

plot(sal_lm2)

#Homoscedasticity seems fine, again residuals normality looks pretty good for majority of data (upper tail is a little strange)

plot(sal_lm3)

#Homoscedasticity seems fine, again residuals normality looks pretty good for majority of data (upper tail is a little strange)

plot(sal_lm4)

#Homoscedasticity seems fine, again residuals normality looks pretty good for majority of data (upper tail is a little strange)

#For the q-q, upper tail could be influenced that as rank/years get high, salaries start to increase non-linearly
```

```{r, include = FALSE}
#VIF for each model

vif(sal_lm1)

#Years_PhD and Years_Service both have VIF values > 4 (7.5, 5.9) - remove one of them from the model? Remove both - is this incorporated into rank?

vif(sal_lm2)
vif(sal_lm3)
vif(sal_lm4)

#No other VIF values were higher than 4. Also, with the exception of Years_PhD in model 2, all were very close to one (and in that instance, the value was 2)

#AIC models comparison

lm1_aic <- AIC(sal_lm1) #9093.82
lm2_aic <- AIC(sal_lm2) #9096.81
lm3_aic <- AIC(sal_lm3) #9097.21
lm4_aic <- AIC(sal_lm4) #9095.45
```

```{r, include = FALSE}
#What about trying two models with interaction terms 

sal_lm5 <- lm(Salary ~ Rank + Sex + Discipline + Years_Service + Discipline*Years_Service, data = faculty_salaries)

summary(sal_lm5)


```

```{r, include = FALSE}
#What about taking out rank and leaving in one of the years?

sal_lm6 <- lm(Salary ~ Discipline + Sex + Years_Service, data = faculty_salaries)

summary(sal_lm6)

```

```{r, include = FALSE}
#I think it's nicer to have assistant professor be the reference level

faculty_salaries$Rank <- factor(faculty_salaries$Rank)

faculty_salaries$Rank <- fct_relevel(faculty_salaries$Rank, "AsstProf")


sal_lm4 <- lm(Salary ~ Rank + Discipline + Sex, data = faculty_salaries)

summary(sal_lm4)

```

Multiple linear regression was used to model faculty salaries, and it was found that a model incorporating Rank, Discipline, and Sex explains a significant amount of variance in salary (Salary ($) = 13723(Associate Professor) + 47403(Professor) + 13709(Applied Discipline) + 4492(Male), F(4,392) = 79.18, p < 0.001, R^2 = 0.45). Three different levels of rank are incorporated into the model: assistant professor, associate professor, and professor; the reference level in the model is assistant professor. On average, associate professors can be expected to make 13,723 more than assistant professors if other factors are held constant, with professors making the most at 47,403 more than assistant professors. Both theoretical(A) and applied(B) disciplines were considered for the model, with the reference level set to theoretical disciplines. On average, it appears as though faculty members working within applied disciplines will make 13,709 more than their colleagues in theoretical disciplines. Lastly, the model looks at the sex of the faculty, with the reference level of the model set to female. Male faculty make 4,492 more than females on average, with all other factors held constant. 


```{r stargazer, results = 'asis', header = FALSE, echo = FALSE}
#stargazer table

stargazer(sal_lm4, 
          type = 'html', 
          digits = 3,
          style = "aer",
          title = "Table X: Faculty Salary Regression Results", 
          dep.var.labels = c("Salary"),
          covariate.labels = c("Assoc. Professor", "Professor", "Applied Discipline", "Male"))

```
<br>

Both rank and discipline appear to be significant drivers of the salary model (p values < 0.001). While the sex of the faculty members does affect average expected salary, it has less of a significant impact on the overall model. Future iterations of the model could include considerations for faculty research area, public or private universities, and size of the department or program. 