---
title: "Part 1"
author: "Claire Madden"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
#Load packages

library(tidyverse)

#Load data

enroll <- read_csv("enrollment final edited.csv")

```

```{r, include=FALSE}
#Visualize the trends in the raw data

enroll_graph <- ggplot(enroll, aes(x = Year)) +
  geom_line(aes(y = Males, color = "Males")) +
  geom_line(aes(y = Females, color = "Females"))
  
enroll_graph  
  
  
```

```{r, include = FALSE}
#Regression models for each

lm_males <- lm(Males ~ Year, data = enroll)

summary(lm_males)

lm_females <- lm(Females ~ Year, data = enroll)

summary(lm_females)

```

```{r, include = FALSE}
#Pearson's R

male_cor = cor.test(enroll$Year, enroll$Males)
male_cor

female_cor = cor.test(enroll$Year, enroll$Females)
female_cor

```


```{r, include=FALSE}
#Look at residuals for each

plot(lm_males)

plot(lm_females)

```

```{r, echo=FALSE, warning = FALSE}
#Plot both models

enroll_model_graph <- ggplot(enroll, aes(x = Year)) +
  geom_point(aes(y = Males, color = "Males")) +
  geom_smooth(aes(y = Males), method = lm, se = TRUE, size = 0.5, color = "gray20") +
  geom_point(aes(y = Females, color = "Females")) +
  geom_smooth(aes(y = Females), method = lm, se = TRUE, size = 0.5, color = "gray20") +
  theme_classic() +
  theme(legend.justification=c(1,0), legend.position=c(0.25,0.7)) +
  scale_x_continuous(expand = c(0,0), limits = c(1967, 2016)) +
  labs(x = "Year", y = "Graduate Enrollment")

enroll_model_graph + labs(colour = "Gender")


```

**Figure X. Relationship between graduate enrollment and year for males and females.** Year significantly predicts total graduate enrollment for both males (b = 9069, t(47) = 16.6, p < 0.001) and females (b = 30,130, t(47) = 51.7, p < 0.001). The models explain a significant amount of variance in graduate enrollment for males (F(1,47) = 276, p < 0.001, R2 = 0.85) and females (F(1,47) = 2669, p < 0.001, R2 = 0.98). There is a strong correlation between year and enrollment for both males (Pearson's R = 0.92) and females (Pearson's R = 0.99) Gray region is the 95% confidence interval for the mean predicted value.