---
title: "Part 3"
author: "Claire Madden"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE, message=FALSE}
# load packages


library(tidyverse)
library(ggplot2)
library(reshape2)
library(dplyr)
library(RColorBrewer)
library(effsize)

# read in data

median_salary <- read_csv("Median salary for doctoral recipients.csv")

```

```{r, include = FALSE}
# exploration of data

# male employed
hist_male_emp <- ggplot(median_salary, aes(x = male_emp))+
  geom_histogram(bins = 10)

hist_male_emp

qq_male_emp <- ggplot(median_salary, aes(sample = male_emp))+
  geom_qq()

qq_male_emp

# female employed
hist_female_emp <- ggplot(median_salary, aes(x = female_emp))+
  geom_histogram(bins = 10)

hist_female_emp

qq_female_emp <- ggplot(median_salary, aes(sample = female_emp))+
  geom_qq()

qq_female_emp


# male post doc

hist_male_pd <- ggplot(median_salary, aes(x = male_pd))+
  geom_histogram(bins = 10)

hist_male_pd

qq_male_pd <- ggplot(median_salary, aes(sample = male_pd))+
  geom_qq()

qq_male_pd

# female post doc

hist_female_pd <- ggplot(median_salary, aes(x = female_pd))+
  geom_histogram(bins = 10)

hist_female_pd

qq_female_pd <- ggplot(median_salary, aes(sample = female_pd))+
  geom_qq()

qq_female_pd

# these are not looking to be very normally distributed
# should i use Mann - Whitney U (wilcoxon rank sum) test for rank? - risk loss of resolution

```






```{r, include = FALSE, message=FALSE, warning=FALSE}
# Does median salary differ significantly between male and female starting postdoc positions?

test_pd <- wilcox.test(median_salary$male_pd, median_salary$female_pd, paired = TRUE)

test_pd

# test received a warning message "cannot compute exact pvalue with ties"

# non significant p value

# cliff's delta

cd_pd <- cliff.delta(median_salary$male_pd, median_salary$female_pd)

cd_pd

# cliff's delta = 0.04 (negligible)

```



```{r, include = FALSE, message=FALSE, warning=FALSE}
# Does median salary differ significantly between male and female PhD recipients in non-postdoc employment positions?

test_emp <- wilcox.test(median_salary$male_emp, median_salary$female_emp, paired = TRUE)

test_emp


# test received a warning message "cannot compute exact pvalue with ties"
# need to find out why test is defaulting to continuity correction

# significant p value = 0.00257 

# cliff's delta

cd_emp <- cliff.delta(median_salary$male_emp, median_salary$female_emp)

cd_emp


# cliff's delta = 0.213 (small)

```




```{r,include = FALSE, message=FALSE, warning=FALSE}

data_melt <- melt(median_salary[c('field', 'male_pd', 'female_pd', 'male_emp', 'female_emp')], id.vars = 1)

```


```{r, echo=FALSE}
# make side by side column graphs of median salary for each field in post doc

# maybe make a new dataframe first


pd_data <- data_melt %>% 
  filter(variable == "female_pd" | variable == "male_pd")

pd_graph <- ggplot(pd_data, aes(x = field, y = value))+
  geom_col(aes(fill = variable), position = "dodge")+
  theme_classic()+
  scale_fill_manual(values = c("plum4", "slategray3"),
                    name = "Group",
                    labels = c("Male Postdoctoral Study", "Female Postdoctoral Study"
                    ))+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,70000),
                     breaks = seq(0,70000, by = 10000))+ 
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 5))+
  labs(x = "Field of Study", y = "Median Salary ($)")
  

pd_graph

# QUESTIONS/OPINIONS
# do we need the legend or can we explain in the figure caption? #I could go either way on the caption.
# should i try to shorten the names of the fields or is the angle enough? #AP here: I think the Angle looks good!  
# 
```

**Figure X. Median Salary for Males and Females in Postdoctoral Study by Field.** Distribution of median salary for males and females working in positions of post doctoratal study seperated by discipline in 2015. There is no significant difference between male and female median salaries within each field of study [W=19.5, p=0.89] (cliff's delta = 0.04 (negligible)). 


```{r, echo=FALSE}
# make side by side column graphs of median salary for each field in employed


emp_data <- data_melt %>% 
  filter(variable == "female_emp" | variable == "male_emp")


emp_graph <- ggplot(emp_data, aes(x = field, y = value))+
  geom_col(aes(fill = variable), position = "dodge")+
  scale_fill_manual(values = c("plum4", "slategray3"),
                    name = "Group",
                    labels = c("Male Postdoctoral Employment", "Female Postdoctoral Employment"))+
  theme_classic()+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,125000),
                     breaks = seq(0,125000, by = 20000))+ 
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 5))+
  labs(x = "Field of Study", y = "Median Salary ($)")

emp_graph

```

**Figure X. Median Salary for Males and Females in Postdoctoral Employment by Field.** Distribution of median salary for males and females working in post doctoratal employment seperated by discipline in 2015. There is no significant difference between male and female median salaries within each field of study [W=101, p=0.0026] (cliff's delta = 0.213 (small)). 



```{r, echo=FALSE}
# add box plot with each column

colnames(data_melt) <- c("Field", "GenderEmployment", "Salary")



box_plot <- ggplot(data_melt, aes(x = GenderEmployment, y = Salary, fill = GenderEmployment))+
  geom_boxplot(show.legend = FALSE)+
  theme_bw()+
  scale_fill_manual(values = c("lightblue3", "lightblue4", "plum3", "plum4"))+
  labs(y = "Median Salary Across Fields ($)", x = "Group")+
  scale_y_continuous(expand = c(0,0),
                     limits = c(30000,125000))+
  scale_x_discrete(labels = c("male_pd" = "Male \nPost Doc \nStudy",
                              "female_pd" = "Female \nPost Doc \nStudy",
                              "male_emp" = "Male \n Post Doc \nEmployment",
                              "female_emp" = "Female \n Post Doc \nEmployment"))
  

box_plot



```

**Figure X. Median Salary by Sex and Postdoctoral Plans.** Distribution of median salary by sex and post doctoral plans for all fields. *what else??? no statistical test info to include on this one*



