---
title: "Part 3"
author: "Claire Madden"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo = FALSE, message=FALSE}
# load packages


library(tidyverse)
library(ggplot2)
library(reshape2)
library(dplyr)
library(RColorBrewer)
library(effsize)

# read in data

median_salary <- read_csv("Median salary for doctoral recipients.csv")

```

```{r, include = FALSE}
# exploration of data

# male employed
hist_male_emp <- ggplot(median_salary, aes(x = male_emp))+
  geom_histogram(bins = 10)

hist_male_emp

qq_male_emp <- ggplot(median_salary, aes(sample = male_emp))+
  geom_qq()

qq_male_emp

# female employed
hist_female_emp <- ggplot(median_salary, aes(x = female_emp))+
  geom_histogram(bins = 10)

hist_female_emp

qq_female_emp <- ggplot(median_salary, aes(sample = female_emp))+
  geom_qq()

qq_female_emp


# male post doc

hist_male_pd <- ggplot(median_salary, aes(x = male_pd))+
  geom_histogram(bins = 10)

hist_male_pd

qq_male_pd <- ggplot(median_salary, aes(sample = male_pd))+
  geom_qq()

qq_male_pd

# female post doc

hist_female_pd <- ggplot(median_salary, aes(x = female_pd))+
  geom_histogram(bins = 10)

hist_female_pd

qq_female_pd <- ggplot(median_salary, aes(sample = female_pd))+
  geom_qq()

qq_female_pd

# these are not looking to be very normally distributed
# should i use Mann - Whitney U (wilcoxon rank sum) test for rank? - risk loss of resolution

```






```{r, include = FALSE, message=FALSE, warning=FALSE}
# Does median salary differ significantly between male and female starting postdoc positions?

test_pd <- wilcox.test(median_salary$male_pd, median_salary$female_pd, paired = TRUE)

test_pd

# test received a warning message "cannot compute exact pvalue with ties"

# non significant p value

# cliff's delta

cd_pd <- cliff.delta(median_salary$male_pd, median_salary$female_pd)

cd_pd

# cliff's delta = 0.04 (negligible)

```



```{r, include = FALSE, message=FALSE, warning=FALSE}
# Does median salary differ significantly between male and female PhD recipients in non-postdoc employment positions?

test_emp <- wilcox.test(median_salary$male_emp, median_salary$female_emp, paired = TRUE)

test_emp


# test received a warning message "cannot compute exact pvalue with ties"
# need to find out why test is defaulting to continuity correction

# significant p value = 0.00257 

# cliff's delta

cd_emp <- cliff.delta(median_salary$male_emp, median_salary$female_emp)

cd_emp


# cliff's delta = 0.213 (small)
# r output says small, my notes say maybe closer to medium

```




```{r,include = FALSE, message=FALSE, warning=FALSE}

data_melt <- melt(median_salary[c('field', 'male_pd', 'female_pd', 'male_emp', 'female_emp')], id.vars = 1)

```


```{r, echo=FALSE}
# make side by side column graphs of median salary for each field in post doc

# maybe make a new dataframe first


pd_data <- data_melt %>% 
  filter(variable == "female_pd" | variable == "male_pd")

pd_graph <- ggplot(pd_data, aes(x = field, y = value))+
  geom_col(aes(fill = variable), position = "dodge")+
  theme_classic()+
  scale_fill_manual(values = c("plum4", "lightblue3"),
                    name = "Group",
                    labels = c("Male", "Female"
                    ))+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,70000),
                     breaks = seq(0,70000, by = 10000))+ 
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 5))+
  labs(x = "Field of Study", y = "Median Salary ($)")
  

pd_graph

# QUESTIONS/OPINIONS
# do we need the legend or can we explain in the figure caption? #I could go either way on the caption. #BG - Maybe shorten the fields in the legend to just Gender - Male/Female like in Part 1?

# should i try to shorten the names of the fields or is the angle enough? #AP here: I think the Angle looks good!  BG - I'm good with the angle unless we really need to find extra space in our report! Then we could shorten the long ones to free up some space

# 
```

**Figure X. Median Salary for Males and Females in Postdoctoral Study by Field.** Distribution of median salary for males and females working in positions of post doctoratal study seperated by discipline in 2015. There is no significant difference between male and female median salaries within each field of study [W=19.5, p=0.89] (cliff's delta = 0.04). Data used with permission from the National Science Foundation. 

Male and female median salaries are equal in Biological and Biomedical Sciences, Chemistry, Economics, Engineering, Geoscience, Atmospheric and Ocean Sciences, Humanities and Arts, and Psychology. Among the remaining fields of study, male median salaries are higher than female median salaries in only Education, Mathematics and Computer Science, and Other non-S&E Fields. The greatest difference in median salaries between males and females is within the category of Other non-S&E Fields, where male median income is $6,000 greater than female median income. 

```{r, echo=FALSE}
# make side by side column graphs of median salary for each field in employed


emp_data <- data_melt %>% 
  filter(variable == "female_emp" | variable == "male_emp")


emp_graph <- ggplot(emp_data, aes(x = field, y = value))+
  geom_col(aes(fill = variable), position = "dodge")+
  scale_fill_manual(values = c("plum4", "lightblue3"),
                    name = "Group",
                    labels = c("Male", "Female"))+
  theme_classic()+
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,125000),
                     breaks = seq(0,125000, by = 20000))+ 
  theme(axis.text.x = element_text(angle = 55, hjust = 1, size = 5))+
  labs(x = "Field of Study", y = "Median Salary ($)")

emp_graph

```

**Figure X. Median Salary for Males and Females in Postdoctoral Employment by Field.** Distribution of median salary for males and females working in post doctoratal employment seperated by discipline in 2015. There is a significant difference between male and female median salaries within each field of study [W=101, p=0.0026] (cliff's delta = 0.213). Data used with permission from the National Science Foundation. 


Within the population of doctoral recipients that are employed, male median salaries are equal to or greater than female median salaries in every field except Physics and Astronomy. The largest difference in median salaries between males and females is in Mathematics and Computer Sciences where men earn a median salary of \$105,000 and women earn a median salary of \$90,000, a difference of \$15,000. 


